\input{macros.tex}

\author{
    Marc \textsc{Chevalier}\\
    Thomas \textsc{Pellissier} \textsc{Tanon}}
\date{\today}
\title{\textsc{Friedman}'s translation}

\begin{document}

\maketitle

\section{\textsc{Friedman}'s Translation}

\begin{definition}
    Let $\R$ be a formula. The \textbf{parametrized negation} is 
    $$
        \ngr := A \Rightarrow \R
    $$
\end{definition}

We gather here some basic properties of the parametrized negation

\begin{proposition}
    In intuitionisctic logic,
    \begin{enumerate}[(i)]
        \item $B \Rightarrow \ngr A \vdash A \Rightarrow \ngr B$
        \item $A\vdash \ngr\ngr A$
        \item $A\Rightarrow B \vdash \ngr B \Rightarrow \ngr A$
        \item $A\Rightarrow B \vdash \ngr\ngr A \Rightarrow \ngr \ngr B$
        \item $\ngr \ngr \ngr A \vdash \ngr A$
    \end{enumerate}
\end{proposition}
\begin{proof}
    \input{src/prop1-2.tex}
\end{proof}

We now define the parametrized translation.
\begin{definition}
    Let $\R$ be a formula. The \textbf{parametrized negative translation} $A^\ngr$ is defined by induction on $A$ as follows:
    \begin{center}
    \begin{tabular}{cccc}
        $\bottom^\ngr := \R$ & $\top^\ngr := \top$ & \multicolumn{2}{c}{$(a\dot{=}b)^\ngr := \ngr\ngr(a\dot{=}b)$}\\
        \multicolumn{2}{c}{$(A\wedge B)^{\neg_R}:=A^\ngr \wedge B^\ngr$} & \multicolumn{2}{c}{$(A \Rightarrow B)^{\neg_R}:=A^\ngr \Rightarrow B^\ngr$}\\
        \multicolumn{4}{c}{$(A \vee B)^{\neg_R}:=\ngr\ngr(A^\ngr \vee B^\ngr)$}\\
        \multicolumn{2}{c}{$\forall x A)^\ngr := \forall x A^\ngr$} & \multicolumn{2}{c}{$\exists x A^\ngr := \ngr\ngr(\exists x A^\ngr)$}
    \end{tabular}
    \end{center}
\end{definition}

Note that $(\neg A)^\ngr = \ngr A^\ngr$. We gather here the basic properties of the parametrized translation.

\begin{proposition}
    In intuitionistic logic,
    \begin{enumerate}[(i)]
        \item $\vdash (A\vee \neg A)^\ngr$
        \item $\R \vdash A^\ngr$
        \item $\ngr\ngr A^\ngr \vdash A^\ngr$
    \end{enumerate}
\end{proposition}
\begin{proof}
    \input{src/prop1-4.tex}
\end{proof}

\begin{theoreme}
    If $\vdash A$ is derivable in classical predicate logic and if no free variable of \R\, occurs in the derivation, then $\vdash A^{\neg_\R}$ is derivable in intuitionisctic predicate logic.
\end{theoreme}
\begin{proof}

\end{proof}

\begin{theoreme}
    If $PA \vdash A$ and if no free variable of \R\, occurs in the derivation, then $HA \vdash A^{\neg_\R}$.
\end{theoreme}
\begin{proof}
    \input{src/th1-6.tex}
\end{proof}

\begin{theoreme}
    If $PA\vdash \forall x,\exists y:(a\dot{=}b)$ then $HA\vdash \forall x, \exists y: (a\dot{=}b)$.
\end{theoreme}
\begin{proof}
    Let us write $F \forall x,G$ where $G$ is a $\Sigma_1^0$ formula. By using $(\forall E)$, if $F$ if provable with PA, so $G$ too. As $G$ is $\sigma_1^0$, $G$ is provable with $HA$. By using $(\forall I)$, we deduce that $F$ is provable with HA too.
    %cf. http://www.lsv.ens-cachan.fr/~goubault/types005.html
\end{proof}

\end{document}
